---
title: "Quick Guide to SEEK-VEC"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dpi=300,
  fig.width = 4,
  fig.height = 4
)
```

## Getting started

Install and load the R package `seekvec` from the Github page [rdanning/seekvec](https://github.com/rdanning/seekvec) and load the package for the topic modeling algorithm of your choice (for this example we will use the [TopicScore](https://cran.r-project.org/web/packages/TopicScore/index.html) package).

```{r install, eval = FALSE}
library(devtools)
library(usethis)

install_github("rdanning/seekvec")
```
```{r load}
library(seekvec)
library(TopicScore)
```

## Load and clean example data

For this tutorial, we will use the [MADStat-Text](https://github.com/ZhengTracyKe/MADStat-Text/blob/main/MADStaText/TextCorpusFinal.RData) dataset from Ke et al. [1] The dataset contains the corpus matrix corresponding to a collection of 83,331 statistical abstracts and a vocabulary of 2,106 words. Detailed information about the dataset can be found [here](https://www.tracyke.net/software/MADStat/ReadMe-for-MADStatPart2.pdf).

```{r load-data-hidden, echo = FALSE}
load("~/Downloads/TextCorpusFinal.RData")
```

```{r load-data, eval = FALSE}
load("TextCorpusFinal.RData")
```

We remove abstracts that do not contain any of the vocabulary words, leaving 81,425 abstracts.

```{r clean-data, warning=FALSE}
D <- as.matrix(TDM)
D <- D[, colSums(D) > 0]
vocabulary <- dimnames(D)$Terms
```

## Run SEEK-VEC

Suppose we wish to consider models with $K=\{3,...,10\}$. We choose a threshold of 10 top words per topic and run SEEK-VEC as follows:

```{r run-seek, warning=FALSE}
SEEK.out <- run_SEEK(D, 3:10, method = "ts", threshold = 10)
```

## Analyze output

We can plot a heatmap to view the structure of the ensemble matrix. We choose a filter value of 0.5, which represents a weighted-majority consensus of a relationship between two words.

```{r plot-seek}
plot_results(SEEK.out, vocabulary, filter.value = 0.5, font.size = 4)
```

We can also view the words with the strongest signal:

```{r top-words-df}
signal.df <- extract_diagonals(SEEK.out, vocabulary)
signal.df[rev(order(signal.df$value))[1:10],]
```

Suppose we examine the candidate topic models and decide that the model with $K=7$ is the most interpretable. We can examine the stability of these derived topics with respect to the ensemble matrix:

```{r stability}
K7.stability <- assess_stability(SEEK.out, vocabulary, K = 7, n.top.words = 5, font.size = 2)
K7.stability$plot
K7.stability$block.stability
```

As shown by the heatmap and the stability metrics, only one of these topics is highly robust to the choice of $K$. 

## References
1. Ke, Z. T., Ji, P., Jin, J. & Li, W. Recent Advances in Text Analysis. Annual Review of Statistics and Its Application 11, 347–372. http://dx.doi.org/10.1146/annurev-statistics-040522-022138 (Apr. 2024).