---
title: "Quick Guide to SEEK-VFI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

set.seed(78)
```

## Getting started

Install and load the R package `seekvfi` from the Github page [rdanning/seekvfi](https://github.com/rdanning/seekvfi).

```{r install, eval = FALSE}
library(devtools)
library(usethis)

install_github("rdanning/seekvfi")
```
```{r load, warning = FALSE, message=FALSE}
library(seekvfi)
library(tidyverse)
```

## Load example data

For this tutorial, we will use the fibroblast reprogramming [dataset](https://zenodo.org/records/1443566) from Treutlein et al. [1] The dataset contains the expression counts of 3379 genes measured on 355 cells. Detailed information about the dataset can be found [here](https://www.nature.com/articles/nature18323).

```{r load-data-hidden, echo = FALSE}
data <- readRDS(paste0("~/Desktop/Trajectory methods/Datasets/fibroblast-reprogramming_treutlein.rds"))
```

```{r load-data, eval = FALSE}
load("fibroblast-reprogramming_treutlein.rds")
```
```{r clean-data}
counts <- t(data$counts)
cell.labels <- data$cell_info$milestone_id
cell.labels <- factor(cell.labels, levels =
                        c("MEF",
                          "d2_intermediate",
                          "d2_induced",
                          "d5_intermediate",
                          "d5_earlyiN",
                          "Neuron",
                          "Myocyte")
                      )
```

## Run SEEK-VFI

Suppose we wish to consider trajectories with $K=\{3,...,12\}$ nodes. We run SEEK-VFI as follows:

```{r run-seek, warning=FALSE, message=FALSE}
seek.scores <- run_seekvfi(counts, 3:12)
```

## Analyze output

We can compare the PCA embeddings of the data using all genes compared to using the top 1000 SEEK-VFI genes:

```{r run-plots, warning=FALSE}
expression <- as.matrix(data$expression)

all.pca.df <- prcomp(t(expression))$rotation[,1:2] %>%
  as.data.frame()
all.pca.df$cls <- cell.labels

seek.genes <- (seek.scores %>% arrange(desc(SV.score)))$gene[1:1000]
seek.idx <- colnames(expression) %in% seek.genes

seek.pca.df <- prcomp(t(expression[,seek.idx]))$rotation[,1:2] %>%
  as.data.frame()
seek.pca.df$cls <- cell.labels

```

```{r plots-all, fig.width=6, fig.height=5}
ggplot(all.pca.df, aes(x = PC1, y = PC2, color = cls)) +
  geom_point(size = 2) +
  theme_bw() +
  xlab("PC1") +
  ylab("PC2") +
  theme(legend.title = element_blank()) +
  ggtitle("All genes")
```

```{r plots-seek, fig.width=6, fig.height=5}
ggplot(seek.pca.df, aes(x = PC1, y = PC2, color = cls)) +
  geom_point(size = 2) +
  theme_bw() +
  xlab("PC1") +
  ylab("PC2") +
  theme(legend.title = element_blank()) +
  ggtitle("Top 1000 SEEK-VFI genes")
```

Restricting to the top 1000 SEEK-VFI genes allows the PCA embedding to more clearly demonstrate the differentiation of the precursor cells into mature neurons and myoctes.

## References
1. Treutlein, B. et al. Dissecting direct reprogramming from fibroblast to neuron using single-cell RNA-seq. Nature 534, 391–395. issn: 1476-4687. http://dx.doi.org/10.1038/nature18323 (June 2016).
